Statistical Machine Learning Models on A Bank Marketing dataset

### 2025-12-08 Slides.Rmd
slides update4

---
title: "Statistical Machine Learning Models on A Bank Marketing dataset"
author: "Sandra, Tigist, Mussa Hassen, Waddah"
subtitle: 'Colorado State University, DSCI 445: Statistical Machine Learning, Dr.Andee
  Kaplan'
output:
  ioslides_presentation:
    widescreen: true
---

@@

--- 
```{r, echo=FALSE}
library(ggplot2)

@@

* **Most clients say NO; only ≈11% say YES**

@@

## Methodology: Data Preparation (Slide 5) (s)

@@

## LogReg: Feature Selection & Linearity  #MUSSA
* Initial feature selection was performed using **Backward Stepwise Elimination** on the full model to find the most useful set of main effects.
* Our selected formula resutled:

$y \sim job + marital + education + balance + housing + $

$loan + contact + day + month + duration + $

$campaign + previous+ poutcome$
* **Residual Plots** were used to assess the linearity assumption for numeric predictors (`age`, `balance`). This led to testing a model with non-linear terms.

@@

## LogReg: Regularization & Tuning   #MUSSA
* We explored Ridge and LASSO regularization as an alternative approach for feature selection.
* **5-Fold Cross-Validation** was used to select the optimal penalty parameter, $\lambda$, which minimized the mean CV error for each regularized model.

```{r, echo=FALSE, results='asis', fig.width=4, fig.height=3}

@@

## LogReg: Evaluation & Coefficients  #MUSSA
* The final selected LogReg model was evaluated on the 5-fold CV error rate, along with metrics focused on the confusion matrix.
* Models performed relatively similar. We would prefer a model with higher **sensitivity** (few false negatives)

@@

---
# 2. K-Nearest Neighbors (KNN) Model
---

@@

**Test Accuracy (k=5):** `r round(knn_accuracy_demo * 100, 2)`%
---

@@

**Optimal k:** `r best_k`
--- 
## KNN 3/3: Evaluation & Test Performance
* The final KNN model (with optimal $k$) was applied to the **Test Set**. We evaluate performance using **AUC** and compare it to the LogReg models.
* KNN generally provides a good intermediate balance—often capturing non-linearity better than LogReg but being less computationally expensive than boosting.

@@

---
# 3. Boosting Model (High Predictive Benchmark)
---
```{r boosting-setup, echo=FALSE, message=FALSE, warning=FALSE}
library(gbm)
library(pROC)
library(caret)

>>>>

---
title: "Statistical Machine Learning Models on A Bank Marketing dataset"
author: "Sandra, Tigist, Mussa Hassen, Waddah"
subtitle: 'Colorado State University, DSCI 445: Statistical Machine Learning, Dr. Andee Kaplan'
output:
  ioslides_presentation: 
    widescreen: true
---

@@

```{r, echo=FALSE}
library(ggplot2)

@@

* **Most clients say NO; only about 11% say YES **

@@

## Methodology: Data Preparation (s)

@@

## LogReg: Feature Selection & Linearity  
* Initial feature selection was performed using **Backward Stepwise Elimination** on the full model to find the most useful set of main effects.
* Our selected formula resutled:

$y \sim job + marital + education + balance + housing + loan +$

$contact + day + month + duration + campaign + previous+ poutcome$

* **Residual Plots** were used to assess the linearity assumption for numeric predictors (`age`, `balance`). This led to testing a model with non-linear terms.

@@

## LogReg: Regularization & Tuning  
* We explored Ridge and LASSO regularization as an alternative approach for feature selection.
* **5-Fold Cross-Validation** was used to select the optimal penalty parameter, $\lambda$, which minimized the mean CV error for each regularized model.

```{r, echo=FALSE, results='asis', fig.width=5, fig.height=3.5}

@@

## LogReg: Evaluation & Coefficients
* The final selected LogReg model was evaluated on the 5-fold CV error rate, along with metrics focused on the confusion matrix.
* Models performed relatively similar. We would prefer a model with higher **sensitivity** (few false negatives)

@@

# 2. K-Nearest Neighbors (KNN) Model

@@

**Test Accuracy (k=5):** `r round(knn_accuracy_demo * 100, 2)`%

@@

**Optimal k:** `r best_k`

## KNN 3/3: Evaluation & Test Performance
* The final KNN model (with optimal $k$) was applied to the **Test Set**. We evaluate performance using **AUC** and compare it to the LogReg models.
* KNN generally provides a good intermediate balance—often capturing non-linearity better than LogReg but being less computationally expensive than boosting.

@@

# 3. Boosting Model (High Predictive Benchmark)

```{r boosting-setup, echo=FALSE, message=FALSE, warning=FALSE}
library(gbm)
library(pROC)
library(caret)
